---
title: "Text Mining 1"
author: "Anthony Kamau"
toc: yes
output: pdf_document
---

---


```{r, message=FALSE}
# install.packages(c("tidytext","tm", "wordcloud", "SnowballC"))

library(tidyverse) 
library(tidytext)  
library(tm)        
library(wordcloud) 
library(SnowballC) 
```

# Introduction
We use the `computer.txt` dataset. It's a data set with reviews of computers. 
This data set is annotated using aspect-based sentiment analysis. Aspect-based 
sentiment analysis is a text analysis technique that categorizes data by specific
aspects and identifies the sentiment attributed to each one. 

## Load data
```{r}
computer_531 <- read_delim(
  file = "computer.txt", 
  delim = "\n", 
  col_names = "review",
  show_col_types = FALSE
)
```
```{r}
head(computer_531)
```

# Preparation and Pre-processing

## Visualization 

### Word cloud
We shall visualize using a word cloud.
```{r, warning=FALSE}
wordcloud(
  words = computer_531$review,
  min.freq = 5,
  max.words = 50,
  random.order = FALSE,
  colors = brewer.pal(8, "Dark2"))

# The dominant terms are monitor, the, acer, screen, price, computer
# From the word cloud we can see that This dataset is dominated by monitor/screen
# reviews (likely Acer laptops/netbooks),
# However, very common function words such as `the`, `and` and `this` are still
# cluttering the picture because stop words haven't been removed yet.
```
### Frequency plot
Another way to visualize the term frequencies is using a barplot. For this we need
to convert the data to a tidy format (where each word is a different row) and 
count the number of occurrences of each word.

```{r}
# tokenize texts
comp_words <- computer_531 |> 
  unnest_tokens(word, review)

# check the resulting tokens
head(comp_words)

# unnest_tokens function from tidytext package breaks the text into individual 
# tokens (a process called tokenization)
```


```{r}
comp_words |> 
  # count the frequency of each word
  count(word) |> 
  # arrange the words by its frequency in descending order
  arrange(desc(n)) |> 
  # select the top 30 most frequent words
  head(30) |> 
  # make a bar plot (reorder words by their frequencies)
  ggplot(aes(x = n, y = reorder(word, n))) + 
  geom_col(fill="gray") +
  labs(x = "frequency", y="words") + 
  theme_minimal()

# This selects the most frequent 30 tokens and plot a bar chat
# We see that many of the top words are `stop words`. 
# These stop words need to be removed
```

### Remove stop words
```{r}
comp_words_no_stop <- 
  comp_words |> 
  # remove stop words
  anti_join(stop_words)

# stop_words is a dataset with stop words. By using anti_join, you keep only the
# words in the first dataset that are not in the stop_words dataset.
```

```{r}
comp_words_no_stop |> 
  # count the frequency of each word
  count(word) |> 
  # arrange the words by its frequency in descending order
  arrange(desc(n)) |> 
  # select the top 30 most frequent words
  head(30) |> 
  # make a bar plot (reorder words by their frequencies)
  ggplot(aes(x = n, y = reorder(word, n))) + 
  geom_col(fill="gray") +
  labs(x = "frequency", y="words") + 
  theme_minimal()
```

### Adjust stopwords list. 
```{r}
df_stopwords <- 
  stop_words |> 
  filter(!word %in% c("order", "ordered", "ordering", "orders")) |> 
  rbind(tibble(word = c("buy", "online"), lexicon = c("Me", "Me")))

# remove stop words
comp_words_no_stop_modified <- 
  comp_words |>
  anti_join(df_stopwords)
```

```{r}
comp_words_no_stop_modified |> 
  # count the frequency of each word
  count(word) |> 
  # arrange the words by its frequency in descending order
  arrange(desc(n)) |> 
  # select the top 30 most frequent words
  head(30) |> 
  # make a bar plot (reorder words by their frequencies)
  ggplot(aes(x = n, y = reorder(word, n))) + 
  geom_col(fill="gray") +
  labs(x = "frequency", y="words") + 
  theme_minimal()

# The resulting barplot looks almost identical to the previous one, just missing
# "buy," confirming the small custom lexicon barely changed the top-30. Most of
# the signal was already captured
```
### Stemming
When we deal with text, often documents contain different versions of one base word,
called a stem. `SnowballC` is a package based on the famous Porter’s stemming 
algorithm that collapses words into their stem. We shall use the `getStemLanguages` 
function to find out how many languages the package supports.

```{r}
getStemLanguages()

# supports 26 languages 
```

We apply stemming and show the 30 most frequent stemmed words
```{r}
comp_words_stemming <- 
  comp_words_no_stop_modified |> 
  mutate(stem = wordStem(word))

comp_words_stemming |> 
  # count the frequency of each word
  count(stem) |> 
  # arrange the words by its frequency in descending order
  arrange(desc(n)) |> 
  # select the top 30 most frequent words
  head(30) |> 
  # make a bar plot (reorder words by their frequencies)
  ggplot(aes(x = n, y = reorder(stem, n))) + 
  geom_col(fill="gray") +
  labs(x = "frequency", y="words") + 
  theme_minimal()

# The resulting top terms (monitor, price, acer, screen... etc) are largely the
# same as before stemming here, since the top words were already in single-form
```

# Regular Expressions
We use regular expressions (regex) to find the reviews in computer_531, which contain
words “Monitor” or “monitor”, “memory” or “Memory”, and “Delivery” or “delivery” and
we see how many reviews contain each pair of words.
```{r}
# reviews containing "Monitor" or "monitor"
reviews_mon <- computer_531 |> 
  filter(str_detect(review, "[Mm]onitor"))

# reviews containing "memory" or "Memory"
reviews_mem <- computer_531 |> 
  filter(str_detect(review, "[Mm]emory"))

# reviews containing "Delivery" or "delivery"
reviews_del <- computer_531 |> 
  filter(str_detect(review, "[Dd]elivery"))

# compare the occurrence of each pair of words
kwords    <- c("Monitor", "Memory", "Delivery")
nr_kwords <- c(nrow(reviews_mon), 
               nrow(reviews_mem), 
               nrow(reviews_del))

tibble(kwords, nr_kwords)

# We use `str_detect` function from `stringr` package to find the presence of 
# patterns of your interest and the `filter function` to only retain the ones
# that contain the patterns.
# Out of the 531 reviews, 105 mention `Monitor`, 5 mention `Memory` and 
# 3 mention `Delivery`
# This is inline with the word cloud and frequency plots which suggested that 
# monitors are the dominant topic, while memory and delivery are barely discussed.
```

```{r}
## Detect
head(str_detect(computer_531$review, 
                "(\\w+) ([Mm]onitor)"))

# Returns TRUE/FALSE per review, telling us whether that review contains a word
# immediately followed by "monitor" anywhere in it. 
# It is useful for filtering or counting, not for seeing what matched.
```

```{r}
## Extract
head(str_extract(computer_531$review, 
                 "(\\w+) ([Mm]onitor)"))

# Pulls out only the substring that matched the whole pattern (both capture groups combined),
# e.g. "this monitor"
# Rows with no match return NA. 
# Only the first match per string is returned by default
# str_extract_all would give all matches
```

```{r}
## Subset
head(str_subset(computer_531$review, 
                "(\\w+) ([Mm]onitor)"))

# Instead of returning TRUE/FALSE or just the matched fragment, this returns the 
# entire original review text for every review that matched.
# essentially review[str_detect(review, pattern)]. 
# It's a shortcut for filtering a character vector down to just the matching elements.
```
```{r}
## Match (Useful with capture groups)
head(str_match(computer_531$review, 
               "(\\w+) ([Mm]onitor)"))

# This is the most detailed of the four. Column 1 is the full match (same as str_extract),
# and columns 2 and 3 are the two capture groups broken out separately. So we get the
# preceding word and "monitor" as two separate values. 
# This is handy when we specifically want to isolate a captured sub-piece 
# (e.g. just the word before "monitor") rather than the whole match.
```

```{r}
# use str_extract_all
str_extract_all(computer_531$review[1:10], 
                '\\b[A-Z]+\\b', 
                simplify = TRUE)

# all matches, returned as a matrix
# Unlike plain str_extract (first match only), str_extract_all returns 
# every match per string. 
# With simplify = TRUE, instead of returning a list (where each element could have
# a different length), it pads all results into a rectangular matrix: each row is
# one of the first 10 reviews, each column is one match position, and empty
# positions (where a review had fewer all-caps words than the row with the most)
# are filled with "". So a row like "I" "" "" means that review had only one 
# all-caps word ("I"), while "A" "I" "" means it had two.
```

```{r}
str_extract(computer_531[[1]][1:20], 
            "^.*##")

# extracts the characters at the beginning of each line until ## for only 
# the first 20 reviews
```

```{r}
computer_531 <- computer_531 |> 
  mutate(cleaned_review = str_extract(review,"##.*$"),
         aspect_sentiment = str_extract(review,"^.*##"))

# check the newly added columns
head(computer_531[-1])

# new column added to the computer_531 data frame with the name cleaned_review,
# which contains only the review text. And another column added with the 
# name aspect_sentiment, which contains the asepcts and sentiment words.
```

# Parsing aspect tags
For an even cleaner version you can use regex's lookbehind and lookahead features
to drop the `##` parts
```{r}
computer_531 <- computer_531 |> 
  mutate(cleaned_review = str_extract(review,"(?<=##).*$"),
         aspect_sentiment = str_extract(review,"^.*(?=##)"))

# check the newly added columns
head(computer_531[-1])

# No leftover `##` characters cluttering either column
```

# Sentiment Scoring
```{r}
# define sum_list function which adds the scores
sum_list <- function(list_values) {
  sum(as.numeric(list_values))
}

computer_531 <- computer_531 |>
  # sentiment_score: extract the scores and sum them using the sum_list function
  mutate(sentiment_score = map_dbl(str_extract_all(aspect_sentiment, "[-|\\+]\\d+"), sum_list),
                               # assign negative when score < 0
         sentiment = case_when(sentiment_score < 0 ~ "negative",
                               # assign negative when score > 0
                               sentiment_score > 0 ~ "positive",
                               # assign neutral otherwise
                               TRUE ~ "neutral"))

head(computer_531)

#  This create a new column sentiment with the values positive, negative and neutral. 
# The code assigns neutral in case when there is no aspect in the corresponding 
# column or the sum of scores is equal to zero, and it assigns negative or positive
# when the sum score is lower or higher) than zero repectively.
```

```{r}
computer_531 |> 
  arrange(sentiment_score) |>
  tail(1) |>
  pull(cleaned_review) |>
  strwrap() |>
  writeLines()

# arrange(sentiment_score) sorts the whole table by score, ascending (most negative 
# first, most positive last)

#tail(1) grabs the last row — i.e., the single review with the highest sentiment score.

# pull(cleaned_review) extracts just the cleaned_review text column from that 
#row (rather than the whole row/data frame).

# The result is the actual text of the most positive review in the dataset
```

